Reanalysis Step 3: Propensity Score Analysis with Twang

### 2025-08-19 reanalysis-3-propensity_scores.qmd
more cleanup

for (i in 1:n_imputations) {
  cat("Processing imputation", i, "of", n_imputations, "...")

  # Run twang ps model - following analysis.Rmd approach
  twang_model <- ps(
    formula = ps_formula,
    data = data.frame(imputed_datasets[[i]]),  # Convert to data.frame as in analysis.Rmd
    n.trees = n_trees,
    interaction.depth = 3,
    shrinkage = 0.01,
    estimand = "ATT",
    stop.method = c("es.mean"),  # Using es.mean only for speed
    verbose = FALSE
  )

@@

  cat("  - Trees used (es.mean):", twang_model$desc$es.mean$n.trees, "")
  cat("  - Weight range: [", round(min(weights), 3), ",", round(max(weights), 3), "]")
}

cat("Twang propensity score calculation complete")
```

@@

  if (i <= max_imputations_for_diagnostics) {
    cat("Balance summary for imputation", i)
    print(summary(twang_model))
  }
}

@@

@@

cat("Generating twang diagnostic plots for", n_to_show, "imputations...")

# For each selected imputation, show twang-specific diagnostics
for (i in imps_to_show) {
  cat("Diagnostics for Imputation", i, ":")

@@

  tryCatch({
    plot(twang_model)
  }, error = function(e) {
    cat("  Standard plots not available, generating custom diagnostics...")

    ps_scores <- twang_results[[i]]$ps_scores
    weights <- twang_results[[i]]$weights
    treated <- imputed_datasets[[i]]$ever_lapse_binary == 1

    par(mfrow = c(2, 2))

    hist(ps_scores[!treated], main = "PS Distribution - Control", 
         xlab = "Propensity Score", col = "lightblue")
    hist(ps_scores[treated], main = "PS Distribution - Treated", 
         xlab = "Propensity Score", col = "lightcoral")

    hist(weights[!treated], main = "Weights - Control", 
         xlab = "Weight", col = "lightblue")
    hist(weights[treated], main = "Weights - Treated", 
         xlab = "Weight", col = "lightcoral")

    par(mfrow = c(1, 1))
  })

  print(summary(twang_model))
}
```

@@

# Plot PS distribution for first few imputations
for (i in imps_to_show) {
  cat("Propensity score distributions for imputation", i, "")

@@

  cat("  Weight summary:")
  cat("    Control: mean =", round(mean(weights[!treated]), 3), 
      ", max =", round(max(weights[!treated]), 3), "")
  cat("    Treated: mean =", round(mean(weights[treated]), 3),
      "(should be 1.0 for ATT)")
}
```

>>>>

# Track progress
progress_df <- data.frame(
  Imputation = integer(),
  Trees_Used = integer(),
  Weight_Min = numeric(),
  Weight_Max = numeric()
)

for (i in 1:n_imputations) {
  # Run twang ps model - following analysis.Rmd approach
  twang_model <- ps(
    formula = ps_formula,
    data = data.frame(imputed_datasets[[i]]),  # Convert to data.frame as in analysis.Rmd
    n.trees = n_trees,
    interaction.depth = 3,
    shrinkage = 0.01,
    estimand = "ATT",
    stop.method = c("es.mean"),  # Using es.mean only for speed
    verbose = FALSE
  )

@@

  # Track progress
  progress_df <- rbind(progress_df, data.frame(
    Imputation = i,
    Trees_Used = twang_model$desc$es.mean$n.trees,
    Weight_Min = round(min(weights), 3),
    Weight_Max = round(max(weights), 3)
  ))
}

# Display progress summary
kable(progress_df, caption = "Twang Model Fitting Progress")
```

@@

  if (i <= max_imputations_for_diagnostics) {
    # Balance summary will be shown in diagnostics section
  }
}

@@

Showing diagnostic plots for the first **`r min(max_imputations_for_diagnostics, n_imputations)`** imputations.

@@

# Generate diagnostic plots for selected imputations
for (i in imps_to_show) {

@@

  # Print section header
  cat("\n### Imputation", i, "\n\n")

  # Display twang diagnostic plots
  plot(twang_model)

  # Display balance summary
  print(summary(twang_model))
}
```

@@

# Create weight distribution summary
weight_dist_summary <- data.frame(
  Imputation = integer(),
  Control_Mean = numeric(),
  Control_Max = numeric(),
  Treated_Mean = numeric()
)

# Plot PS distribution for first few imputations
for (i in imps_to_show) {

@@

  # Store weight summary
  weight_dist_summary <- rbind(weight_dist_summary, data.frame(
    Imputation = i,
    Control_Mean = round(mean(weights[!treated]), 3),
    Control_Max = round(max(weights[!treated]), 3),
    Treated_Mean = round(mean(weights[treated]), 3)
  ))
}

# Display weight distribution summary
kable(weight_dist_summary, 
      caption = "Weight Distribution by Imputation",
      col.names = c("Imputation", "Control Mean", "Control Max", "Treated Mean (should be 1.0 for ATT)"))
```
